\documentclass[10pt, conference, compsocconf,letter]{IEEEtran}   	
%\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
						% TeX will automatically convert eps --> pdf in pdflatex

\usepackage{color}
\newcommand{\fix}[1]{\textcolor{red}{#1}}
\newcommand{\highlight}[1]{\textcolor{red}{#1}}
\usepackage{amssymb, amsmath}
\usepackage{mathtools}
\usepackage{booktabs}
%\usepackage[noend]{algpseudocode}
%\usepackage[ruled,vlined]{algorithm2e}
\usepackage{verbatim}
\newcommand*\rot{\rotatebox{90}}
%\usepackage[compatible]{algpseudocode}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{varwidth}
\usepackage{multirow}
\usepackage{rotating, xcolor}
\usepackage{subfig}
\usepackage{enumitem}
%\usepackage[caption=false]{subfig}
\usepackage[hyphens]{url}
\usepackage{float}
\newfloat{algorithm}{t}{lop}

\usepackage{tcolorbox}
\definecolor{mycolor}{rgb}{0.122, 0.435, 0.698}
%\begin{tcolorbox}[width=\linewidth, boxsep=0pt, left=-4pt, boxrule=0.5pt]
%\end{tcolorbox} 
%\newtcolorbox{mybox}{colback=red!5!white,colframe=mycolor}
\makeatletter
\newcommand{\mybox}[1]{%
  %\setbox0=\hbox{#1}%
  %\setlength{\@tempdima}{\dimexpr\wd0+13pt}%
  \begin{tcolorbox}[colframe=mycolor,boxrule=0.5pt,arc=4pt,
      left=-4pt,right=6pt,top=6pt,bottom=6pt,boxsep=0pt,width=\linewidth]
    #1
  \end{tcolorbox}
}
\makeatother

\algnewcommand{\LineComment}[1]{\State \(\triangleright\) #1}
%\newcommand{\procdecl}[1]   {\proc{#1}\vrule width0pt height0pt depth 7pt \relax}
\newcommand{\lilabel}[1]        {\label{li:#1}}

\newcommand{\erdosrenyi}{Erd\H os-R\'{e}nyi }
\newcommand{\qg}{\u{g}}
\newcommand{\qG}{\u{G}}
\newcommand{\qc}{\c{c}}
\newcommand{\qC}{\c{C}}
\newcommand{\qs}{\c{s}}
\newcommand{\qS}{\c{S}}
\newcommand{\qu}{\"{u}}
\newcommand{\qU}{\"{U}}
\newcommand{\qo}{\"{o}}
\newcommand{\qO}{\"{O}}
\newcommand{\qI}{\.{I}}
\newcommand{\wa}{\^{a}}
\newcommand{\wA}{\^{A}}
\newcommand{\xor}{\mbox{xor}}
\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\renewcommand{\floatpagefraction}{.9}
\newcommand{\minusone}{\text{-}1}

%% ABAB: Shortcuts/macros below are not used but it can be referenced as a cheatsheet
\newcommand{\liref}[1]      {line~\ref{li:#1}}
\newcommand{\Liref}[1]      {Line~\ref{li:#1}}
\newcommand{\lirefs}[2]     {lines \ref{li:#1}--\ref{li:#2}}
\newcommand{\Lirefs}[2]     {Lines \ref{li:#1}--\ref{li:#2}}
\newcommand{\lireftwo}[2]   {lines \ref{li:#1} and~\ref{li:#2}}
\newcommand{\lirefthree}[3] {lines \ref{li:#1}, \ref{li:#2}, and~\ref{li:#3}}

\def\Cpp{C{}\texttt{++}}
\newcommand{\mA}{\mathbf{A}} 
\newcommand{\mL}{\mathbf{L}}
\newcommand{\mU}{\mathbf{U}}
\newcommand{\transpose}     {^{\mbox{\scriptsize \sf T}}}
\newcommand{\mB}{\mathbf{B}}
\newcommand{\mC}{\mathbf{C}}
\newcommand{\dimN}{n}
\newcommand{\dimM}{m}
\newcommand{\dimK}{k}
\newcommand{\dnzc}{\id{nzc}}
\newcommand{\dnzr}{\id{nzr}}
\newcommand{\dni}{\id{ni}}
\newcommand{\dnnz}{\id{nnz}}
\newcommand{\dsort}{\id{sort}}
\newcommand{\dscan}{\id{scan}}
\newcommand{\dsearch}{\id{search}}
\newcommand{\dmin}{\func{min}}
\newcommand{\dmax}{\func{max}}
\newcommand{\dth}{th}
\newcommand{\dlen}{\id{len}}
\newcommand{\matlab}{{\sc MATLAB}}

%
%TGM: included and new commands from our GraphBLAS C API document
%
\usepackage{listings}
\renewcommand{\vector}[1]{{\bf #1}}
\renewcommand{\matrix}[1]{{\bf #1}}
\renewcommand{\arg}[1]{{\sf #1}}
\newcommand{\zip}{{\mbox{zip}}}
\newcommand{\zap}{{\mbox{zap}}}
\newcommand{\ewiseadd}{{\mbox{\bf ewiseadd}}}
\newcommand{\ewisemult}{{\mbox{\bf ewisemult}}}
\newcommand{\mxm}{{\mbox{\bf mxm}}}
\newcommand{\vxm}{{\mbox{\bf vxm}}}
\newcommand{\mxv}{{\mbox{\bf mxv}}}
\newcommand{\gpit}[1]{{\sf #1}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\nan}{{\sf NaN}}
\newcommand{\nil}{{\bf nil}}
\newcommand{\ifif}{{\bf if}}
\newcommand{\ifthen}{{\bf then}}
\newcommand{\ifelse}{{\bf else}}
\newcommand{\ifendif}{{\bf endif}}
\newcommand{\zero}{{\bf 0}}
\newcommand{\one}{{\bf 1}}
\newcommand{\true}{{\sf true}}
\newcommand{\false}{{\sf false}}
\newcommand{\syntax}{{C Syntax}}
%%%%%%

\hyphenation{Suite-Sparse}
\hyphenation{Graph-BLAS}
\hyphenation{Suite-Sparse-Graph-BLAS}

%\input{algobox}

\title{LAGraph: A Community Effort to Collect Graph Algorithms built on top of the GraphBLAS}


\author{
Tim Mattson$^\ddag$, Tim Davis$^\diamond$, Manoj Kumar$^\P$, Ayd\i n Bulu\qc$^\dag$, Scott McMillan$^\S$, Jos\'e Moreira$^\P$, Carl Yang$^{*,\dag}$ \\
\\

 \normalsize
 {\em $^\ddag$Intel Corporation}  {\em $^\dag$Computational Research Division, Lawrence Berkeley National Laboratory} \\
{\em $^\diamond$Texas A and M University} ~~ {\em $\P$IBM Corporation}~~ {\em $\S$Software Engineering Institute, Carnegie Mellon University}   \\
{\em $^*$Electrical and Computer Engineering Department, University of California, Davis}
 }



\begin{document}
\maketitle

\begin{abstract}

In 2013, we released a position paper to launch a community effort to define a 
common set of building blocks for constructing graph algorithms ``in the language of Linear Algebra''.
This led to the GraphBLAS.   We released a specification for the C programming language binding to the
GraphBLAS in 2017.  Since that release, multiple libraries that conform to the 
GraphBLAS C specification have been produced.

In this position paper, we launch the next phase of this ongoing community effort; a project to
assemble a set of high level graph algorithms built on top of the GraphBLAS.  While many 
of these algorithms are well-known with high quality implementations available,
they have not been assembled in one place and integrated with the GraphBLAS.  We call 
this project the LAGraph graph algorithm project and with this position paper, put out a 
call for collaborators to join us.  While the initial goal is to just assemble
these algorithms, we hope in the long run this effort will result production worthy code;
with the LAGraph library serving as an open source repository of verified graph algorithms that use the
GraphBLAS. 

\end{abstract}

\begin{IEEEkeywords}
Graph Algorithms, Linear Algebra, GraphBLAS
\end{IEEEkeywords}

\input{intro}

\input{GraphBLAS}

\input{LAGraphStruct}

\input{disc}

\input{conclusion}

\section*{Acknowledgments}

Tim Davis would like to acknowledge the support of SuiteSparse:GraphBLAS by MIT
Lincoln Laboratory, Intel, and the National Science Foundation (NSF
CNS-1514406).

TODO: Ayd\i n Bulu\qc\, Carl Yang and maybe Scott will need to acknowledge someone.

\bibliographystyle{IEEEtran}
\bibliography{LAGraph19}

\end{document}  
