
\subsection{SuiteSparse:GraphBLAS}

SuiteSparse:GraphBLAS is a full implementation of the GraphBLAS standard, which
defines a set of sparse matrix operations on an extended algebra of semirings
using an almost unlimited variety of operators and types.  When applied to
sparse adjacency matrices, these algebraic operations are equivalent to
computations on graphs.  GraphBLAS provides a powerful and expressive framework
for creating graph algorithms based on the elegant mathematics of sparse matrix
operations on a semiring.

The design of a GraphBLAS library is flexible, because its data structures are
opaque to the user.  SuiteSparse:GraphBLAS uses a compressed-sparse vector
storage mechanism, in four different forms.  A matrix can be stored in
row-major order (CSR), or column-major order (CSC).  Each sparse vector
consists of a sorted list of indices, and the corresponding numerical values.
The sparse vectors are packed together into two arrays, and another ``pointer
array'' (of size equal to the dimension of the matrix, say $n$) keeps track of
where each row (or column) starts.  The space taken is $O(n+e)$ for a CSR
matrix with $n$ rows or a CSC matrix with $n$ columns, and with $e$ entries.
Both of these two forms can be modified in a {\em hypersparse} form
\cite{BulucGilbert08}, where the pointer array itself becomes sparse, and
non-empty vectors take no space at all.  The space is reduced to $O(e)$,
however, so that matrices with enormous dimensions can be created, as long as
$e << n$.  Hypersparsity is exploited automatically.  All methods can operate
on all four matrix formats in any combination.

The ability to incrementally modify a graph is critical in many applications.
GraphBLAS includes two operations that can make small incremental changes to a
graph/matrix:  namely \verb'GrB_setElement' and \verb'GrB_assign', and It would
be exceedingly slow to insert or delete a single new entry in a CSR or CSC
format, taking $O(n+e)$ time {\bf per entry} inserted.  Instead, the
non-blocking idea in GraphBLAS is exploited.  Fast deletion of entries is
handled by creating {\em zombies}, which are entries tagged for later deletion.
Fast insertion is handled with {\em pending tuples}, which is a separate
unordered list of $(i,j,a_{ij})$ for each new entry.  When a matrix operation
occurs (such as matrix multiply), all zombies are killed and all pending tuples
are assembled, in a single $O(n+ e + p \log p)$ step (for $p$ pending tuples),
or $O(e +p \log p)$ in the hypersparse case.  As a result, it is just as fast
to use a sequence of $e$ \verb'GrB_setElement' operations to build a matrix, as
it is to create an array of $e$ tuples and use \verb'GrB_build'.  Internally,
SuiteSparse:GraphBLAS is building the list itself, for the user, and then does
a \verb'GrB_build' when the matrix needs to be completed.

To enable high-performance matrix-matrix multiply, a code generation mechanism
is used to build functions for each semiring that can be created with built-in
operators.  The functions can rely on Gustavson's method \cite{Gustavson78}, a
dot product method, and heap-based method, all with masked variants.  A current
prototype of the package adds an early exit mechanism for the MIN, MAX, OR, and
AND monoids, where a dot product can terminate as soon as a terminal value is
found in the result (\verb'true' for OR, for example).  This will enable a fast
direct-optimizing BFS to be written, where the ``pull'' is a dot product, and
the ``push'' a saxpy-based operation (Gustavson's or the heap method).

Since it is meant as the GraphBLAS reference implementation, testing is a vital
component to the package.  In SuiteSparse:GraphBLAS, each GraphBLAS operation
was written twice: once in high-performance algorithms in C, and again in a
very simple and short MATLAB script, using dense matrices with of the required
type.  For example, \verb'GrB_assign' requires about 3,908 lines of C (not
counting comments), but only 161 lines in MATLAB.  The pattern in the MATLAB
version is held as a seperate boolean matrix.  The MATLAB functions are not
intended to be fast.  Instead, they exactly mimic the GraphBLAS API
Specification, line by line, so they can be visually inspected for conformance
to the spec.  For example, matrix multiply is written with a triply-nested
\verb'for' loop.  Then, to test the package, each computation is done in
SuiteSparse:GraphBLAS (via a MATLAB interface) and in the MATLAB mimic.  The
tests pass only if the results are identical in both value and pattern (even
with identical floating-point roundoff error, in most cases).

SuiteSparse:GraphBLAS appears in Debian and Ubuntu linux distros,
and has been released as part of the RedisGraph database
module of the Redis database systems, by RedisLabs, Inc.
\url{https://redislabs.com/redis-enterprise/redis-modules/redis-enterprise-modules/redisgraph/}

